---
title: "wrangle"
author: "Fournier Raphaël"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file wrangles the pilot data for the familiarity face perception study.

## load the libraries that we will be using ## 

## install ##

```{r install-pkg}
#install.packages(c("tidyverse", "RColorBrewer", "patchwork", "gtsummary"))
```

take a snapshot of loaded packages and update the lock.file using renv

```{r snapshot-renv}
# take a snapshot and update the lock.file
renv::snapshot() # this is only necessary when new packages or installed or packages are updated.
```

## load ##

```{r load-pkg}
pkg <- c("tidyverse", "RColorBrewer", "patchwork", "gtsummary", "paletteer", "here")

lapply(pkg, library, character.only = TRUE)

## Set the amount of dodge in figures
pd <- position_dodge(0.7)
pd2 <- position_dodge(1)

## Theme settings for ggplot
theme_set(
  theme_bw() +
    theme(text = element_text(size = 20), 
          title = element_text(size = 22),
          legend.position = "bottom")
)

#Colors
myColors <- c("Novel" = paletteer_d("rcartocolor::Safe",4)[1],"Learned" =  paletteer_d("rcartocolor::Safe",4)[2])
```



# 1. Read in behavioural data

```{r}
# Set the file path

#Without blocks!!!
raw <- list.files(
    path = here("data/behavioural_data"),
    pattern = "*csv",
    full.names = TRUE,
    recursive = TRUE
  ) %>%
  set_names() %>%
  map_dfr(read_csv, col_types = cols(Gender = col_character()))
```

#Saving raw data
```{r}
write_csv(raw, here("data", "raw_behav_data_WP3.csv"))

raw <- read_csv(here("data", "raw_behav_data_WP3.csv"))
```



# 2. Wrangle: face recognition task

```{r}
#Wrangle the raw data to extract the face recognition task data
n_blocks      <- 3
trials_per_block <- 60   # 180 trials total per participant

data <- raw %>%
  filter(!is.na(trial_in_block)) %>%
  rename(trial = recog_trial.thisN,cond = familiarity, acc = key_resp_5.corr, rt = key_resp_5.rt, pid = participant) %>%
  mutate(trial = trial+1) %>%
  relocate(pid, .before = cond) %>%
  select(pid, trial,acc, cond, rt, block) %>%
  relocate(any_of(c("block", "cond")), .before = acc) %>%
  arrange(pid) %>%
  mutate(cond = fct_recode(cond, Novel = "unfamiliar", Learned = "learned"))

```  

---

# 3. Exclusions: face recognition task

Each stage below tracks how many participants are removed and why.

```{r}
#| label: exclusion-tracking

# Helper: count remaining pids
n_pids <- function(df) length(unique(df$pid))

# Running exclusion log — we'll append to this
excl_log <- tibble(
  stage     = character(),
  criterion = character(),
  n_removed = integer(),
  n_remain  = integer()
)

log_exclusion <- function(log, stage, criterion, before, after) {
  bind_rows(log, tibble(
    stage     = stage,
    criterion = criterion,
    n_removed = as.integer(before - after),
    n_remain  = as.integer(after)
  ))
}

n_start <- n_pids(data)
```

## 3.1 Missed trials (> 40% NA rt)

```{r}
#| label: excl-missed
total_trials   <- n_blocks * trials_per_block  # 180
missed_thresh  <- 0.40

missed_summary <- data |>
  group_by(pid) |>
  summarise(n_missed = sum(is.na(rt)), .groups = "drop") |>
  mutate(prop_missed = n_missed / total_trials)

pids_missed <- missed_summary |>
  filter(prop_missed > missed_thresh) |>
  pull(pid)

data_excl1 <- data |> filter(!pid %in% pids_missed)

excl_log <- log_exclusion(
  excl_log, "1", 
  glue::glue("> {missed_thresh * 100}% missed trials"),
  n_start, n_pids(data_excl1)
)

# Visualise missed-trial distribution
missed_trials_plot <- missed_summary |>
  mutate(excluded = pid %in% pids_missed) |>
  ggplot(aes(x = prop_missed, fill = excluded)) +
  geom_histogram(binwidth = 0.02, colour = "white") +
  geom_rug(data = ~filter(., excluded), 
           aes(colour = excluded), sides = "b", linewidth = 1) +
  geom_vline(xintercept = missed_thresh, linetype = "dashed", colour = "red") +
  scale_fill_manual(values = c("FALSE" = "steelblue", "TRUE" = "tomato"),
                    labels = c("Retained", "Excluded")) +
  labs(
    title = "Proportion of missed trials",
    x     = "Proportion missed",
    y     = "Count",
    fill  = NULL
  )
missed_trials_plot
```
## 3.2 RT range violations (> 25% of trials outside 0.2–2.0 s)

```{r}
#| label: excl-rt-range

rt_lo   <- 0.2
rt_hi   <- 2.5
rt_thresh <- 0.25   # > 25% violations → exclude

rt_check <- data_excl1 |>
  filter(!is.na(rt)) |>
  group_by(pid) |>
  summarise(
    n_trials      = n(),
    n_out_of_range = sum(rt < rt_lo | rt > rt_hi),
    prop_out       = n_out_of_range / n_trials,
    .groups        = "drop"
  )

pids_rt <- rt_check |>
  filter(prop_out > rt_thresh) |>
  pull(pid)

data_excl2 <- data_excl1 |> filter(!pid %in% pids_rt)

excl_log <- log_exclusion(
  excl_log, "2",
  glue::glue("> {rt_thresh * 100}% RT outside [{rt_lo}, {rt_hi}] s"),
  n_pids(data_excl1), n_pids(data_excl2)
)

# Visualise RT violation distribution
rt_check_plot <- rt_check |>
  mutate(excluded = pid %in% pids_rt) |>
  ggplot(aes(x = prop_out, fill = excluded)) +
  geom_histogram(binwidth = 0.02, colour = "white") +
  geom_rug(data = ~filter(., excluded), 
           aes(colour = excluded), sides = "b", linewidth = 1) +
  geom_vline(xintercept = rt_thresh, linetype = "dashed", colour = "red") +
  scale_fill_manual(values = c("FALSE" = "steelblue", "TRUE" = "tomato"),
                    labels = c("Retained", "Excluded")) +
  labs(
    title = "Proportion of RTs",
    x     = "Proportion out of range",
    y     = "Count",
    fill  = NULL
  )
rt_check_plot
```
## 3.3 Accuracy floor (< 55%) and ceiling (> 95%)

```{r}
#| label: excl-accuracy

acc_lo <- 55
acc_hi <- 95

# Remove out-of-range RTs before accuracy check (trial-level filter)
data_valid_rt <- data_excl2 |>
  filter(!is.na(rt), rt >= rt_lo, rt <= rt_hi)

acc_check <- data_valid_rt |>
  group_by(pid) |>
  summarise(
    n_trials = n(),
    n_correct = sum(acc),
    perc_acc  = n_correct / n_trials * 100,
    .groups   = "drop"
  )

pids_acc_floor   <- acc_check |> filter(perc_acc < acc_lo) |> pull(pid)
pids_acc_ceiling <- acc_check |> filter(perc_acc > acc_hi) |> pull(pid)
pids_acc         <- union(pids_acc_floor, pids_acc_ceiling)

data_excl3 <- data_valid_rt |> filter(!pid %in% pids_acc)

excl_log <- log_exclusion(
  excl_log, "3a",
  glue::glue("Accuracy < {acc_lo}%"),
  n_pids(data_valid_rt), n_pids(data_valid_rt) - length(pids_acc_floor)
)
excl_log <- log_exclusion(
  excl_log, "3b",
  glue::glue("Accuracy > {acc_hi}%"),
  n_pids(data_valid_rt) - length(pids_acc_floor),
  n_pids(data_excl3)
)

# Visualise accuracy distribution
acc_check_plot <- acc_check |>
  mutate(
    excluded = case_when(
      perc_acc < acc_lo ~ "Floor",
      perc_acc > acc_hi ~ "Ceiling",
      TRUE              ~ "Retained"
    ),
    excluded = factor(excluded, levels = c("Retained", "Floor", "Ceiling"))
  ) |>
  ggplot(aes(x = perc_acc, fill = excluded)) +
  geom_histogram(binwidth = 2, colour = "white") +
  geom_rug(data = ~filter(., excluded %in% c("Floor", "Ceiling")), 
           aes(colour = excluded), sides = "b", linewidth = 0.25) +
  geom_vline(xintercept = c(acc_lo, acc_hi), linetype = "dashed", colour = "red") +
  scale_fill_manual(
    values = c("Retained" = "steelblue", "Floor" = "tomato", "Ceiling" = "orange")
  ) +
  labs(
    title = "Accuracy",
    x     = "Accuracy (%)",
    y     = "Count",
    fill  = NULL
  )
acc_check_plot
```
## 3.4 Behavioural exclusion summary

```{r}
#| label: excl-summary

excl_log |>
  knitr::kable(
    col.names = c("Stage", "Criterion", "N removed", "N remaining"),
    caption   = "Participant exclusions — face recognition task"
  )

cat(
  "\nFinal N (face recognition task):", n_pids(data_excl3), "\n",
  "Started with:", n_start, "\n",
  "Total excluded:", n_start - n_pids(data_excl3), "\n"
)

write_csv(excl_log, "data/exclusions.csv")

## and plot
p_exclusions <- missed_trials_plot | rt_check_plot | acc_check_plot
p_exclusions

ggsave("figures/descriptive/data_exclusions.png",
       width = 15, height = 7)
```


#Exclusion pupil data
```{r}
data_clean <- data_excl3 %>%
  filter(!pid =="003")

  
write_csv(data_clean, "data/data.csv")
```

#Saving demo_data
```{r}
age_demo <- read_csv("data/Master_list.csv") |>
  filter(!is.na(AGE))
demo_WP3 <- raw %>%
  select(participant, `Dominant eye`, Gender, Handedness)
demo_WP3 <- demo_WP3[complete.cases(demo_WP3),]
demo_WP3 <- demo_WP3 %>%
  unique() |>
  mutate(age = age_demo$AGE) |>
  filter(participant %in% data_clean$pid)

write_csv(demo_WP3, here("data","data_demo_WP3.csv"))
```
## section 3 - create summary data ##

first, create summary data for rt and accuracy at the pid level and at the 
group level for three tasks

### rt ###
at the pid level
```{r}
rt_pid <- exp_3_data %>% 
  group_by(pid, block, cond) %>%
  summarise(mean = mean(rt),
            sd = sd(rt),
            n = n(), # n here is the total trials per grouping
            sem = (sd/sqrt(n)),
            ci = sem*1.96)
rt_pid
```

at the group level

```{r}
rt_group <- exp_3_data %>% 
  group_by(block, cond) %>% 
  drop_na(rt) %>%
  summarise(mean = mean(rt),
            sd = sd(rt),
            n = length(unique(pid)), # n here is the total subjs
            sem = (sd/sqrt(n)),
            ci = sem*1.96)
rt_group
```

### acc ###

at the pid level

```{r}
acc_pid <- exp_3_data %>% 
  group_by(pid, block, cond) %>% 
  summarise(sum = sum(acc),
            n = n(), # n here is the total trials per grouping
            perc = sum/n*100)
acc_pid
```

at the group level

```{r}
acc_group <- acc_pid %>% 
  group_by(block, cond) %>% 
  summarise(mean_sum = mean(sum),
            n = n(), # n here is the total trials per grouping
            mean_perc = mean(perc),
            sd=sd(perc),
            sem = (sd/sqrt(n))) %>% 
  rename(sum=mean_sum, perc=mean_perc)
acc_group
```


### Without block ###

## rt group ##
```{r}
rt_group_wblock <- exp_3_data %>% 
  group_by(cond) %>% 
  drop_na(rt) %>%
  summarise(mean = mean(rt),
            sd = sd(rt),
            n = length(unique(pid)), # n here is the total subjs
            sem = (sd/sqrt(n)),
            ci = sem*1.96)
rt_group_wblock
```
## rt pid ##
```{r}
rt_pid_wblock <- exp_3_data %>% 
  group_by(pid,cond) %>% 
  drop_na(rt) %>%
  summarise(mean = mean(rt),
            sd = sd(rt),
            n = length(unique(pid)), # n here is the total subjs
            sem = (sd/sqrt(n)),
            ci = sem*1.96) %>%
  left_join(rt_group_wblock,by = join_by(cond))

rt_pid_wblock
```


## acc group ##
```{r}
acc_group_wblock <- acc_pid %>% 
  group_by(cond) %>% 
  summarise(mean_sum = mean(sum),
            n = n(), # n here is the total trials per grouping
            mean_perc = mean(perc),
            sd=sd(perc),
            sem = (sd/sqrt(n))) %>% 
  rename(sum=mean_sum, perc=mean_perc)
acc_group_wblock
```
## acc pid ##

```{r}
acc_pid_wblock <- acc_pid %>% 
  group_by(pid,cond) %>% 
  summarise(mean_sum = mean(sum),
            n = n(), # n here is the total trials per grouping
            mean_perc = mean(perc),
            sd=sd(perc),
            sem = (sd/sqrt(n))) %>% 
  rename(sum=mean_sum, perc=mean_perc) %>%
  left_join(acc_group_wblock, by = join_by(cond))
acc_pid_wblock
```

## section 4 - plot ##

just a few quick and dirty plots for now


## Face recognition experiment ##

density

rt

```{r}
p4.1 <- ggplot(exp_3_data, aes(x=rt, fill=cond, colour=cond)) +
   geom_density(alpha=0.5) +
   scale_fill_manual(values = myColors) +
   scale_color_manual(values = myColors) +
   facet_grid(~block) +  
   ggtitle("Response time by block and condition") +
   xlab("Response time [s]") +
   ylab("Density") +
   xlim(0,3)
p4.1

ggsave("figures/rt_density_pid.jpeg",
       width = 8, height = 6)
```

acc

```{r}
p4.2 <- ggplot(exp_3_data, aes(x=acc, fill=cond, colour=cond)) +
   geom_density(alpha=0.5) +
   scale_fill_manual(values = myColors) +
   scale_color_manual(values = myColors) +
   ggtitle("Accuracy by block and condition") +
   facet_wrap(~block) +
   xlab("Accuracy") +
   ylab("Density")  
p4.2 

#ggsave("/Users/rfournier/Rstudio/pilot5_dmc/figures_exp_1/acc_density_pid.jpeg",
       #width = 8, height = 6)
```

summary plots

rt

```{r}
p4.3 <- ggplot(rt_pid, aes(x=block, y=mean, 
                           fill=cond, colour=cond)) +
   geom_point(alpha = 3) +
   geom_line(aes(group = cond)) +
   scale_fill_manual(values = myColors) +
   scale_color_manual(values = myColors) +
   ggtitle("Mean response time by block and condition") +
   facet_wrap(~pid) +
   xlab("Blocks") +
   ylab("Mean response time [s]") +
  ylim(0,3)
p4.3

ggsave("figures/rt_line_pid.jpeg",
      width = 15, height = 12)
```

acc

```{r}
p4.4 <- ggplot(acc_pid, aes(x=block, y=perc, 
                           fill=cond, colour=cond)) +
   geom_point(alpha = 3) +
   geom_line(aes(group = cond)) +
   scale_fill_manual(values = myColors) +
   scale_color_manual(values = myColors) +
   ggtitle("Accuracy by block and condition") +
   facet_wrap(~pid) +
   xlab("Blocks") +
   ylab("Accuracy [%]") + ylim(c(0,100))
  
p4.4

ggsave("figures/acc_line_pid.jpeg",
       width = 15, height = 12)
```

group level summary data

rt

```{r}
p4.5 <- ggplot(rt_group, aes(x=block, y=mean, 
                           fill=cond, colour=cond)) +
   geom_point(alpha = 3, position = pd2) +
   geom_line(aes(group = cond), position = pd2) +
   geom_errorbar(aes(ymin = mean-sem, ymax = mean+sem),
                width=.2, position=pd2) +
   scale_fill_manual(values = myColors) +
   scale_color_manual(values = myColors) +
   ggtitle("Response time by block and condition") +
   xlab("Blocks") +
   ylab("Mean response time [s]") +
   ylim(0,2)
p4.5

ggsave("figures/rt_line_group.jpeg",
        width = 8, height = 6)
```

acc

```{r}
p4.6 <- ggplot(acc_group, aes(x=block, y=perc, 
                           fill=cond, colour=cond)) +
   geom_point(alpha = 3, position = pd2) +
   geom_line(aes(group = cond), position = pd2) +
   geom_errorbar(aes(ymin = perc-sem, ymax = perc+sem),
                width=.2, position=pd2) +
   scale_fill_manual(values = myColors) +
   scale_color_manual(values = myColors) +
   ggtitle("Accuracy by block and condition") +
   xlab("Blocks") +
   ylab("Accuracy [%]") +
   ylim(0,100)

p4.6

ggsave("figures/acc_line_group.jpeg",
        width = 8, height = 6)
```

plot group data together

```{r}
p4.7 <- p4.5 | p4.6
p4.7

ggsave("figures/rt_acc_line_pilot.jpeg",
        width = 15, height = 6)
```

barplot rt
```{r}
p.4.11 <- ggplot(rt_pid_wblock, aes(x = cond,y=mean.y, 
                           fill=cond, colour=cond)) +
   geom_bar(stat = "summary") +
   geom_errorbar(aes(ymin = mean.y-sem.y, ymax = mean.y+sem.y), col = "black",
                width=.2, position=pd2) +
  geom_point(aes(y = mean.x),
             colour = "black",
             stat = "identity",
             position=pd2,
             alpha = .7) +
   scale_fill_manual(values = myColors) +
   scale_color_manual(values = myColors) +
   ggtitle("Response time by condition") +
   xlab("Conditions") +
   ylab("Response time [s]") +
   ylim(0,2) +
  theme(legend.position = "none")

p.4.11

ggsave("figures/rt_barplot_group.jpeg",
        width = 6, height = 8)
```


barplot acc
```{r}
p.4.12 <- ggplot(acc_pid_wblock, aes(x = cond,y=perc.y, 
                           fill=cond, colour=cond)) +
   geom_bar(stat = "summary") +
   geom_errorbar(aes(ymin = perc.y-sem.y, ymax = perc.y+sem.y), col = "black",
                width=.2, position=pd2) +
    geom_point(aes(y = perc.x),
             colour = "black",
             stat = "identity",
             position=pd2,
             alpha = .7)+
   scale_fill_manual(values = myColors) +
   scale_color_manual(values = myColors) +
   ggtitle("Accuracy by condition") +
   xlab("Conditions") +
   ylab("Accuracy [%]") +
   ylim(0,100) +
  theme(legend.position = "none")

p.4.12

ggsave("figures/acc_barplot_group.jpeg",
        width = 6, height = 8)
```

Two barplots together 
```{r}
p.4.13 <- p.4.11 | p.4.12
p.4.13

ggsave("figures/acc_rt_barplot_group.jpeg",
        width = 15, height = 10)
```

