---
title: "GAMM_pupil"
author: "Fournier Raphaël"
date: "`r Sys.Date()`"
output: html_document
---

# load the libraries that we will be using #

## install ##

```{r install-pkg}
install.packages("remotes")
remotes::install_github("stan-dev/cmdstanr")

install.packages("devtools")
devtools::install_github("jmgirard/standist")
# 
install.packages(c("tidyverse", "RColorBrewer", "patchwork", "brms",
                   "tidybayes", "bayesplot", "future"))
```

## load ##

```{r load-pkg}
pkg <- c("cmdstanr", "standist", "tidyverse", "RColorBrewer", "patchwork", 
         "brms", "tidybayes", "bayesplot", "future", "parallel", "rstan", 
         "here", "flextable", "modelr","paletteer")

lapply(pkg, library, character.only = TRUE)
```

## settings ##

```{r set-options}
options(brms.backend = "cmdstanr",
        mc.cores = parallel::detectCores(),
        future.fork.enable = TRUE,
        future.rng.onMisuse = "ignore") ## automatically set in RStudio

rstan_options(auto_write = TRUE)

supportsMulticore()

detectCores()
```

## plot settings ##

theme settings for ggplot

```{r, eval = F}
theme_set(
  theme_bw() +
    theme(text = element_text(size = 18, face = "bold"), 
          title = element_text(size = 18, face = "bold"),
          legend.position = "bottom")
)
safe_pal <- paletteer_d("rcartocolor::Safe", 4)

myColors_4 <- c(
  "Novel" = safe_pal[1],
  "Learned" = safe_pal[2]
)
## Set the amount of dodge in figures
pd <- position_dodge(0.7)
pd2 <- position_dodge(1)
```

## table settings ##

```{r}
set_flextable_defaults(
  theme_fun = theme_vanilla,
  background.color = "white"
)
```

```{r}
data <- read_csv("data/clean_pupil_data_binned.csv") 
head(data)
str(data)

datam <- data |>
  filter(time >=0 & time <=2400) |>
  mutate(accuracy = as.factor(accuracy),
         accuracy = fct_recode(accuracy, false = "0", true = "1"),
         condition = as.factor(condition),
         condition = fct_recode(condition, Learned = "learned", Novel = "unfamiliar"),
         int_cond = interaction(condition, accuracy, sep = "."),
         int_cond = as.factor(int_cond))

#save data
write_csv(datam, "data/datam_processed.csv")
```

# section 1 #

## quick plot to check the data look as they should ##

create summary data at the pid level

```{r}
data_pid_acc <- datam %>%
  group_by(pid, time, int_cond) %>% 
  summarise(mean = mean(`Pupil_Diameter_bc.mm`, na.rm = TRUE),
            sd = sd(`Pupil_Diameter_bc.mm`, na.rm = TRUE),
            n = length(unique(trial/2)), # 
            sem = (sd/sqrt(length(unique(trial/2)))),
            ci = sem*1.96)
head(data_pid_acc)
```

create summary data at the group level

```{r}
data_group_acc <- datam %>%
  group_by(int_cond, time) %>% 
  summarise(mean = mean(`Pupil_Diameter_bc.mm`, na.rm = TRUE),
            sd = sd(`Pupil_Diameter_bc.mm`, na.rm = TRUE),
            n = length(unique(pid)), # n here is the total pids per condition
            sem = (sd/sqrt(length(unique(pid)))),
            ci = sem*1.96)
head(data_group_acc)
```

```{r}
data_group_acc_no_index <- datam %>%
  group_by(condition, time, accuracy) %>% 
  summarise(mean = mean(`Pupil_Diameter_bc.mm`, na.rm = TRUE),
            sd = sd(`Pupil_Diameter_bc.mm`, na.rm = TRUE),
            n = length(unique(pid)), # n here is the total pids per condition
            sem = (sd/sqrt(length(unique(pid)))),
            ci = sem*1.96)
head(data_group_acc_no_index)


```
diff-accuracy
```{r}
diff_summary <- data_group_acc_no_index %>%
  select(condition, time, accuracy, mean, sem) %>%
  pivot_wider(names_from = accuracy, values_from = c(mean, sem)) %>%
  mutate(mean_diff = mean_true - mean_false,
         sem_diff = sqrt(sem_true^2 + sem_false^2)) %>%
  select(condition, time, mean_diff, sem_diff)
```



## ribbon plot ##

with sem at the pid level

```{r}
p1.1 <- ggplot(data_pid_acc,
               aes(x = time, y = mean, fill = accuracy)) +
  geom_line(aes(colour = accuracy), alpha = 1, linewidth = 1) +
  geom_ribbon(aes(ymin=mean-sem, ymax=mean+sem), alpha = 0.5) +
  scale_colour_brewer(palette = "Dark2") +
  scale_fill_brewer(palette = "Dark2") +
  scale_x_continuous(breaks=seq(1500,4000,500),
                     labels= seq(0,2.5, 0.5)) +
  scale_y_continuous(limits=c(-0.25, 0.75),
                     breaks = seq(-0.25, 0.75, 0.25)) +
  facet_grid(condition~pid) 
p1.1
```

with sem at the group level

```{r}
p1.2 <- ggplot(data_group_acc,
               aes(x = time, y = mean, fill = int_cond)) +
  geom_line(aes(colour = int_cond), alpha = 1, linewidth = 1) +
  geom_ribbon(aes(ymin=mean-sem, ymax=mean+sem), alpha = 0.5) +
  scale_colour_brewer(palette = "Dark2") +
  scale_x_continuous(breaks=seq(0,2400,400),
                     labels= seq(0,2.4, 0.4)) +
  scale_y_continuous(limits=c(-0.25, 0.5),
                     breaks = seq(-0.25, 0.5, 0.25)) +
  scale_fill_brewer(palette = "Dark2") +
  labs(title = "Mean pupil size")
p1.2

ggsave("figures/plot_mean_intr_cond.jpeg", width = 10, height = 8, dpi = 150)
```


diff_plots
```{r}
p1.2 <- ggplot(diff_summary,
               aes(x = time, y = mean_diff, fill = condition)) +
  geom_line(aes(colour = condition), alpha = 1, linewidth = 1) +
  geom_ribbon(aes(ymin=mean_diff-sem_diff, ymax=mean_diff+sem_diff), alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "solid", colour = "darkred") +
  scale_colour_brewer(palette = "Dark2") +
  scale_x_continuous(breaks=seq(0,2400,400),
                     labels= seq(0,2.4, 0.4)) +
  scale_y_continuous(limits=c(-0.25, 0.25),
                     breaks = seq(-0.25, 0.25, 0.25)) +
  scale_fill_brewer(palette = "Dark2") +
  labs(title = "Pupil size difference between true and false")+
  ylab("Pupil size (mm)")
p1.2

ggsave("figures/plot_mean_diff_acc.jpeg", width = 10, height = 8, dpi = 150)
```





# section 2 #

#Load model
```{r}
bp1 <- readRDS("models/model_interaction.rds")
```


## build some models ##

## b1 - accuracy x condition ##

## index coding ##
```{r}
index_data <- datam %>%
  select(pid, trial, time, `Pupil_Diameter_bc.mm`, condition, accuracy) %>%
  mutate(int_cond = as.factor(interaction(condition,accuracy)) %>%
           relevel(ref = "Novel.true"))
```

## formula ##

```{r}
formula = bf(`Pupil_Diameter_bc.mm` ~ 1 + int_cond +
               s(time, by = int_cond, bs = "bs", k = 10) +
               s(time, pid, bs = "fs", xt = list(bs = "bs")))
```

```{r}
get_prior(formula,
          data = index_data, family = gaussian())
```
## set priors ##

```{r}
priors <- c(
  set_prior("normal(0, 0.25)", class = "Intercept"),
  set_prior("normal(0, 0.25)", class = "b"),
  set_prior("normal(0, 0.25)", class = "sds"),
  set_prior("normal(0, 0.25)", class = "sigma")
)
```

#Model 1
```{r}
plan(multicore)
bp1 <- brm(formula = formula,
        data = index_data, family = gaussian(),
        prior = priors,
        iter = 2000, warmup = 1000, cores = 8, chains = 4,
        save_pars = save_pars(all=TRUE),
        control = list(adapt_delta = 0.99),
        seed = 123,
        file = "models/GAMM_pupil")
summary(bp1)
```

#Just checking autocorrelation
```{r}
resids <- residuals(bp1)[, "Estimate"]

index_data <- index_data |>
  mutate(row_id = row_number())

model_vars <- all.vars(formula(bp1)$formula) 

index_data_used <- index_data |>
  filter(if_all(all_of(model_vars), ~ !is.na(.)))

nrow(index_data_used) == nrow(bp1$data)  # should now match

bp1$data <- bp1$data |>
  bind_cols(index_data_used$trial) |>
  rename(trial = `...5`)

some_trial <- unique(bp1$data$trial)[67]

acf(resids[bp1$data$trial == some_trial], lag.max = 50)
```


#Model 2
```{r}
data <- index_data %>%
  mutate(int_cond = factor(int_cond) %>%
           relevel(ref = "Novel.true"))  # set whichever level you want as baseline

fit_releveled <- update(bp1, newdata = data)  # or refit from scratch

plan(multicore)
bp1 <- brm(formula = formula,
        data = index_data, family = gaussian(),
        prior = priors,
        iter = 2000, warmup = 1000, cores = 8, chains = 4,
        save_pars = save_pars(all=TRUE),
        control = list(adapt_delta = 0.99),
        seed = 123,
        file = "models/model_interaction")
summary(bp1)
```


```{r}
vars <- as_tibble(get_variables(bp1)) %>%
  rename(term=value) %>% 
  filter(! term %in% c("lprior", "accept_stat__", "treedepth__", "lp__",
                       "divergent__", "n_leapfrog__", "energy__",
                       "stepsize__"),
         str_detect(term, "zs", negate = TRUE)) %>% 
  print(n=Inf)
```


#Conditional effects

```{r}
## ---------------------------------------------------------------
## Plot conditional effects for the pupil diameter GAMM
## Model: Pupil_Diameter_bc.mm ~ 1 + int_cond + s(time, by = int_cond) +
##                                s(time, pid, bs = "fs")
## ---------------------------------------------------------------

library(brms)
library(tidyverse)

# ------------------------------------------------------------------
# 1. Load your fitted model (adjust path/object name as needed)
# ------------------------------------------------------------------
# fit <- readRDS("path/to/your_model.rds")

# ------------------------------------------------------------------
# 2. Get conditional effects for the time x condition smooth
#    We explicitly ask for the interaction so each condition
#    gets its own fitted smooth trajectory over time.
# ------------------------------------------------------------------
ce <- conditional_effects(
  fit,
  effects = "time:int_cond",
  re_formula = NA,        # population-level only, ignore participant smooth
  method = "fitted"       # posterior mean + credible interval
)

# Pull out the underlying data frame for custom plotting
ce_df <- as.data.frame(ce[["time:int_cond"]])

# ------------------------------------------------------------------
# 3. Plot: fitted trajectories per condition with credible ribbons
# ------------------------------------------------------------------
p_smooths <- ggplot(ce_df, aes(x = time, y = estimate__,
                                colour = int_cond, fill = int_cond)) +
  geom_ribbon(aes(ymin = lower__, ymax = upper__),
              alpha = 0.15, colour = NA) +
  geom_line(linewidth = 1) +
  labs(
    x = "Time (s)",
    y = "Baseline-corrected pupil diameter (mm)",
    colour = "Condition",
    fill = "Condition",
    title = "Fitted pupil diameter trajectories by condition"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom")

print(p_smooths)

ggsave("figures/conditional_effects_time_by_condition.png", p_smooths,
       width = 8, height = 5.5, dpi = 300)

# ------------------------------------------------------------------
# 4. Optional: split into separate panels (facet) if lines overlap
# ------------------------------------------------------------------
p_facet <- p_smooths + facet_wrap(~ int_cond) + theme(legend.position = "none")
print(p_facet)

ggsave("conditional_effects_time_by_condition_facet.png", p_facet,
       width = 9, height = 6, dpi = 300)

# ------------------------------------------------------------------
# 5. Optional: pairwise difference smooths (e.g. Learned.true vs Novel.true)
#    Useful if you want to see WHEN conditions diverge, not just IF.
#    This uses posterior draws directly rather than conditional_effects().
# ------------------------------------------------------------------
plot_diff_smooth <- function(fit, cond_a, cond_b, time_range = NULL) {

  newdata_grid <- fit$data %>%
    distinct(time) %>%
    arrange(time)

  if (!is.null(time_range)) {
    newdata_grid <- newdata_grid %>%
      filter(time >= time_range[1], time <= time_range[2])
  }

  # brms still validates that formula variables exist in newdata even when
  # re_formula = NA excludes them from the calculation, so we need a
  # placeholder pid column. Any single existing level works — its value
  # doesn't affect the output since the by-participant smooth is excluded.
  dummy_pid <- fit$data$pid[1]

  nd_a <- newdata_grid %>% mutate(int_cond = cond_a, pid = dummy_pid)
  nd_b <- newdata_grid %>% mutate(int_cond = cond_b, pid = dummy_pid)

  draws_a <- fitted(fit, newdata = nd_a, re_formula = NA, summary = FALSE)
  draws_b <- fitted(fit, newdata = nd_b, re_formula = NA, summary = FALSE)

  diff_draws <- draws_a - draws_b

  diff_summary <- tibble(
    time      = newdata_grid$time,
    estimate  = apply(diff_draws, 2, mean),
    lower     = apply(diff_draws, 2, quantile, probs = 0.025),
    upper     = apply(diff_draws, 2, quantile, probs = 0.975)
  )

  ggplot(diff_summary, aes(x = time, y = estimate)) +
    geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
    geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2, fill = "steelblue") +
    geom_line(colour = "steelblue", linewidth = 1) +
    labs(
      x = "Time (s)",
      y = paste0("Difference (", cond_a, " − ", cond_b, ")"),
      title = paste0("Estimated difference smooth: ", cond_a, " vs ", cond_b)
    ) +
    theme_minimal(base_size = 13)
}

# Example usage — adjust condition labels to match your factor levels exactly
p_diff_learned <- plot_diff_smooth(bp1, cond_a = "Learned.true", cond_b = "Learned.false")
p_diff_novel <- plot_diff_smooth(bp1, cond_a = "Novel.true", cond_b = "Novel.false")
p_diff_acc_true <-plot_diff_smooth(bp1, cond_a = "Learned.true", cond_b = "Novel.true")
p_diff_acc_false <- plot_diff_smooth(bp1, cond_a = "Learned.false", cond_b = "Novel.false")

print(p_diff_novel)

ggsave("figures/diff_smooth_Noveltrue_vs_Novelfalse.png", p_diff_novel,
       width = 7, height = 5, dpi = 300)
ggsave("figures/diff_smooth_Learnedtrue_vs_Learnedfalse.png", p_diff_learned,
       width = 7, height = 5, dpi = 300)
ggsave("figures/diff_smooth_Learnedtrue_vs_Noveltrue.png", p_diff_acc_true,
       width = 7, height = 5, dpi = 300)
ggsave("figures/diff_smooth_Learnedfalse_vs_Novelfalse.png", p_diff_acc_false,
       width = 7, height = 5, dpi = 300)
## ---------------------------------------------------------------
## Notes:
## - Check exact factor level names in your data: levels(fit$data$int_cond)
## - re_formula = NA excludes the by-participant smooth s(time, pid), so
##   these are population-level (average) trajectories.
## - If conditional_effects() is slow with 4000 draws, you can subsample:
##   ndraws = 1000 inside conditional_effects()/fitted() calls.
## ---------------------------------------------------------------
```


#Effect size
```{r}
# ---------------------------------------------------------------
## 6. Standardized effect sizes (coefficient / sigma), full posterior
##    Rather than dividing point estimates, we divide draw-by-draw so
##    the resulting standardized effect has its own proper CI.
## ---------------------------------------------------------------
library(ggdist)  # for stat_halfeye
 
draws <- as_draws_df(bp1)
 
# Identify the population-level (b_) coefficient columns for int_cond,
# excluding the intercept and any smooth-related columns.
cond_cols <- draws %>%
  select(starts_with("b_int_cond")) %>%
  colnames()
 
# Draw-level standardized effects: each condition coefficient divided by
# that same draw's residual SD (sigma), giving a full posterior of the
# standardized effect rather than a single point-estimate ratio.
std_effects <- draws %>%
  select(all_of(cond_cols), sigma) %>%
  mutate(across(all_of(cond_cols), ~ .x / sigma, .names = "std_{.col}")) %>%
  select(starts_with("std_")) %>%
  pivot_longer(everything(), names_to = "term", values_to = "std_effect") %>%
  mutate(term = str_remove(term, "^std_b_int_cond"))
 
# Numeric summary: mean, 95% CI of the standardized effect per condition
std_summary <- std_effects %>%
  group_by(term) %>%
  summarise(
    estimate = mean(std_effect),
    lower    = quantile(std_effect, 0.025),
    upper    = quantile(std_effect, 0.975),
    .groups = "drop"
  )
 
print(std_summary)
 
# Visual: posterior distributions of standardized effects, with
# conventional small/medium/large reference lines (0.2 / 0.5 / 0.8)
# shown only as a rough anchor, not a verdict.
p_std <- ggplot(std_effects, aes(x = std_effect, y = term)) +
  geom_vline(xintercept = c(-0.2, 0.2), linetype = "dotted", colour = "grey60") +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
  stat_halfeye(fill = "steelblue", alpha = 0.7) +
  labs(
    x = "Standardized effect (coefficient / sigma)",
    y = NULL,
    title = "Posterior distributions of standardized condition effects",
    subtitle = "Dotted lines mark a conventional 'small effect' (|d| = 0.2) reference"
  ) +
  theme_minimal(base_size = 13)
 
print(p_std)
 
ggsave("figures/standardized_effects_posterior.png", p_std,
       width = 7.5, height = 5, dpi = 300)
 
## ---------------------------------------------------------------
## 7. Optional: variance explained by int_cond (more rigorous than
##    a Cohen's-d-style ratio, since it accounts for the smooth terms
##    too). Requires refitting a reduced model without int_cond, or
##    performance::r2_bayes() if you already have both models saved.
## ---------------------------------------------------------------
# library(performance)
# r2_full    <- r2_bayes(fit)
# r2_reduced <- r2_bayes(fit_no_int_cond)   # fit this without int_cond
# variance_explained_by_int_cond <- r2_full$R2_Bayes - r2_reduced$R2_Bayes
 
## ---------------------------------------------------------------
## Notes:
## - Check exact factor level names in your data: levels(fit$data$int_cond)
## - re_formula = NA excludes the by-participant smooth s(time, pid), so
##   these are population-level (average) trajectories.
## - If conditional_effects() is slow with 4000 draws, you can subsample:
##   ndraws = 1000 inside conditional_effects()/fitted() calls.
## ---------------------------------------------------------------
```


now, let's use add_epred_draws()

by pid

```{r}
epred_pid <- index_data %>%
  group_by(pid, int_cond) %>% 
  data_grid(time = seq(0, 2400, 10)) %>%
  add_epred_draws(bp1) ## default here is re_forumla = NULL, which inlcudes all varying effects of pid
head(epred_pid)
```


summarise epreds at the group level

```{r}
epred_group <- epred_pid %>%
  group_by(int_cond, time, .draw) %>%
  summarise(.epred_group = mean(.epred, na.rm =TRUE),
            sd_group = sd(.epred, na.rm = TRUE))
head(epred_group)
```

quick epred plot

```{r}
p2.4 <-
  ggplot(data = epred_group, aes(x = time, y = .epred_group,
                                colour=int_cond, fill = int_cond)) +
  stat_lineribbon(alpha = 0.5, .width = .95) +
  scale_fill_brewer(palette = "Dark2") +
  scale_colour_brewer(palette = "Dark2") +
  scale_x_continuous(limits=c(0,2400),
                     breaks=seq(0,2400,400),
                     labels=seq(0,2.4,0.4)) +
  scale_y_continuous(limits=c(-0.25, 0.5),
                     breaks = seq(-0.25, 0.5, 0.25)) +
  labs(title = "posterior epreds")
p2.4

ggsave("figures/gamm_model/epred.jpeg")
```
## calculate quantile intervals and make a table ##

quantile intervals at the group level (just pick 1s timebins to make it simpler)

```{r}
pred_group_q <- pred_group %>%
  group_by(.draw) %>%
  slice(seq(1, n()),10) %>%
  mutate(time=factor(time)) %>%
  group_by(int_cond, time) %>% 
  median_qi(.prediction_group)
head(pred_group_q)
```

make a table

```{r}
## make a table using flextable (still needs a title but that can be fixed later, depending on the doc it is being used in)
pred_tab <- pred_group_q %>%
  filter(time %in% seq(0, 2400, by = 100)) %>%
  select(-.width, -.point, -.interval) %>%
  mutate(across(where(is.double), \(x) round(x, 2))) %>% 
  flextable()
pred_tab

## save it
save_as_image(pred_tab, path = "figures/pred_table.png")

## this table still needs polishing, but you get the basic idea

## and some tabular data maybe
write_csv(pred_group_q, "data/pred_group_q.csv")
```

# section 4 - posterior predictions by difference score #

## calculate difference scores based on condition (reward > neutral) ##

wrangle epred for contrasts

```{r}
epred_group_2 <- epred_group |>
  separate(int_cond, into = c("Condition", "Accuracy"), sep = "\\.", remove = FALSE) |>
  mutate(Condition = factor(Condition, levels = c("Learned", "Novel")),
       Accuracy = factor(Accuracy, levels = c("false", "true")))

summ_traces <- epred_group_2 %>%
  group_by(Accuracy, Condition, time) %>%
  summarise(est   = median(.epred_group),
            lower = quantile(.epred_group, .025),
            upper = quantile(.epred_group, .975),
            .groups = "drop")

# 2. Difference Learned - Novel, computed WITHIN each draw, then summarised ---
summ_diff <- epred_group_2 %>%
  ungroup()|>
  select(Accuracy, Condition, time, .draw, .epred_group) %>%
  pivot_wider(names_from = Condition, values_from = .epred_group) %>%
  mutate(diff = Learned - Novel) %>%
  group_by(Accuracy, time) %>%
  summarise(est   = median(diff),
            lower = quantile(diff, .025),
            upper = quantile(diff, .975),
            .groups = "drop")

# Shared y-limits so correct and incorrect rows are directly comparable
trace_lims <- range(c(summ_traces$lower, summ_traces$upper))
diff_lims  <- range(c(summ_diff$lower,  summ_diff$upper, 0))

# 3. One row = trace panel + difference panel ---------------------------------
make_row <- function(acc, label) {

  p_trace <- summ_traces %>%
    filter(Accuracy == acc) %>%
    ggplot(aes(time, est, colour = Condition, fill = Condition)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), alpha = .2, colour = NA) +
    geom_line(linewidth = 1) +
    scale_colour_manual(values = myColors_4) +
    scale_fill_manual(values = myColors_4) +
    coord_cartesian(ylim = trace_lims) +
    labs(title = paste0(label, ": pupil trace"),
         x = "Time", y = "Predicted pupil size") +
    theme_classic()

  p_diff <- summ_diff %>%
    filter(Accuracy == acc) %>%
    ggplot(aes(time, est)) +
    geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
    geom_ribbon(aes(ymin = lower, ymax = upper), fill = "grey40", alpha = .25) +
    geom_line(linewidth = 1) +
    coord_cartesian(ylim = diff_lims) +
    labs(title = paste0(label, ": Learned \u2212 Novel"),
         x = "Time", y = "Difference in pupil size") +
    theme_classic()

  p_trace + p_diff
}

p_correct   <- make_row("true",  "Correct trials")
p_incorrect <- make_row("false", "Incorrect trials")

# 4. Stack: correct on top, incorrect below -----------------------------------
p_all <- (p_correct / p_incorrect) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

p_all

# ggsave("pupil_predictions.png", p_all, width = 10, height = 8, dpi = 300)
```

plot

```{r}
p4.4 <- ggplot(data = epred_diff_condition, aes(x = time, y = .epred_group_acc, color = Accuracy, fill = Accuracy)) +
  stat_lineribbon(.width = .95, alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "solid", colour = "darkred") +
  scale_fill_brewer(palette = "Dark2") +
  scale_colour_brewer(palette = "Dark2") +
  scale_x_continuous(limits=c(0,2400),
                     breaks=seq(0,2400,400),
                     labels=seq(0,2.4,0.4)) +
  scale_y_continuous(limits=c(-0.25, 0.25),
                     breaks = seq(-0.25, 0.25, 0.25)) +
  labs(title = "Pupil trace predictions for Correct and Incorrect trials")
p4.4

ggsave ("figures/epred_diff_acc_group.jpeg",
         width = 10, height = 8, dpi = 800)
```

diff: Novel-Learned
```{r}
p4.5 <- ggplot(epred_diff_condition, aes(x = time, y = diff)) +  
  stat_lineribbon(.width = .95, alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "solid", colour = "darkred") +
  scale_fill_brewer(palette = "Dark2") +
  scale_colour_brewer(palette = "Dark2") +
  scale_x_continuous(limits=c(0,2400),
                     breaks=seq(0,2400,400),
                     labels=seq(0,2.4,0.4)) +
  scale_y_continuous(limits=c(-0.50, 0.25),
                     breaks = seq(-0.50, 0.25, 0.25)) +
  theme(legend.position = "none") +
  labs(title = "preds (Novel > Learned)")
p4.5

ggsave ("figures/epred_diff_cond_group.jpeg",
         width = 10, height = 8, dpi = 800)
```

plot together

pred
```{r}
p4.3 <- (p4.1 | p4.2) +
  plot_layout(axes = "collect")
p4.3

ggsave ("figures/pred_diff.jpeg",
        width = 10, height = 8, dpi = 800)
```
epred
```{r}
p4.6 <- (p4.4 | p4.5) +
  plot_layout(axes = "collect")
p4.6

ggsave ("figures/epred_diff.jpeg",
        width = 10, height = 8, dpi = 800)
```

