---
title: "main_effects"
author: "Fournier Raphaël"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## load packages ##
```{r load-pkg}


#install.packages("remotes")
#remotes::install_github("stan-dev/cmdstanr")
# 
#install.packages("devtools")
#devtools::install_github("jmgirard/standist")
# 
#install.packages(c("EMC2", "tidyverse", "RColorBrewer", "patchwork", "tidybayes", "DataExplorer","ggpubr", "paletteer","corrplot", "brms", "bayesplot", "future", "mgcv", "rstan","marginaleffects", "here"))

pkxg <- c("EMC2", "tidyverse", "RColorBrewer", "patchwork", "tidybayes", "DataExplorer","ggpubr", "paletteer","corrplot", "brms", "bayesplot", "future", "mgcv", "rstan","marginaleffects", "here")

lapply(pkxg, library, character.only = TRUE)
```


## plot settings ##
#Colors
```{r}
myColors <- c("Novel" = paletteer_d("rcartocolor::Safe",4)[1],"Learned" =  paletteer_d("rcartocolor::Safe",4)[2])
```


theme settings for ggplot

```{r, eval = F}
theme_set(
  theme_bw() +
    theme(text = element_text(size = 18, face="bold"), 
          title = element_text(size = 18, face="bold"),
          legend.position = "bottom")
)

## Set the amount of dodge in figures
pd <- position_dodge(0.7)
pd2 <- position_dodge(1)
```


#Load data if needed
```{r}
derived_samples_wide_pupil <- read_csv(here("data", "derived_samples_wide_pupil.csv"))


#Save the correlations draws
correlation_draws_max <- read_csv(here("data", "correlation_draws_max.csv"))
correlation_draws_mean <- read_csv(here("data", "correlation_draws_mean.csv"))
correlation_draws_rate <- read_csv(here("data", "correlation_draws_rate.csv"))
```


#Load models
```{r}
load("models/LBA_WP3.RData")
```


#Preregistered Model

#Parameters per subjects
```{r}
#Preregistered model
param_1_pid = EMC2::parameters(LBA_WP3,selection="alpha",N = NULL)
#Remember:number of iters = 5000 / number of chains = 3 -> 5'000 x 35(subjects) x 3(number of chains) = 525000 obs for each parameter
write_csv(param_1_pid, "data/param_pid.csv")
```


###Preregistered model

# section 2 #

## what about visualising parameters from the model? ##

As well as the posterior predictions, we also want to plot and summarise the estimated parameters from the posterior distribution. e.g., drift rate, response caution etc. So let's take a look at that.

Note to Raph/reminder for Rich: One thing that took a little (or rather a lot of) figuring out was how the DMC compare.p convenience function calculates group average parameter estimates for the posterior distribution. It takes all pid posteriors, averages them first and then calculates the relevant quantiles to display and make inferences from. In other words, it first summarises parameters over pid 1, 2, 3, etc. and then calculates quantiles. This matters, because you could also calculate the quantiles without first averaging. So, the below has been tested to produce the same outputs as the compare.p function, which is re-assuring.

But instead of using compare.p, we take the posterior distributions from the samples object and then summarise and plot the results ourselves.

## wrangle all parameters and take a quick look ##

This is just to get a quick overview of all 9 parameters (before we perform any contrasts/comparisons between conditions).

## wrangle the samples first into a tidy format ##

transpose samples and pluck theta out into a list of arrays

## create a group summary distribution ##

summarise across pid (to be consistent with the DMC compare.p function)

##Wrangle
Pid level
```{r}
split_tibble <- function(tibble, col = 'col') tibble %>% split(., .[, col])
array_theta_1 <- split_tibble(param_1_pid, "subjects")
array_theta_1<-do.call(rbind, array_theta_1)
array_theta_pid_1 <- array_theta_1 %>%
  pivot_longer(cols = c("v_SNovel:lMTRUE", "v_SNovel:lMFALSE", "v_SLearned:lMFALSE", "v_SLearned:lMTRUE", "sv_lMTRUE", "B_lRNovel", "B_lRLearned", "A", "t0"), names_to = "param", values_to = "value")%>%
  add_column(datapoint = rep(1:15000,each=9,times=31))%>% #here 30'000 = number of chains (3) x number of iteration (10'000)
  rename(pid=subjects)
```

Average across participants
```{r}
theta_group_1 <- array_theta_pid_1 %>%
  group_by(datapoint,param) %>% 
  summarise(value = mean(value, na.rm = T)) %>% 
  ungroup() %>% 
  add_column(pid = "average") %>% ## add a column which makes clear it is the avg. not needed but useful for plotting later
  relocate(pid, .before = everything())
head(theta_group_1)
glimpse(theta_group_1)
```

## calculate quantiles ##

calculate quantiles per parameter at the group level

```{r}
library(tidybayes)
theta_group_1_q <- theta_group_1 %>%
  group_by(param) %>% 
  median_qi(value)
theta_group_1_q

write_csv(theta_group_1, "data/theta_group.csv")
write_csv(theta_group_1_q, "data/theta_group_q.csv")
```

and do the same per pid

```{r}
theta_pid_1_q <- array_theta_pid_1 %>%
  group_by(pid, param) %>% 
  median_qi(value)
theta_pid_1_q

write_csv(array_theta_pid_1, "data/theta_pid.csv")
write_csv(theta_pid_1_q, "data/theta_pid_q.csv")
```

## some quick and dirty plots with all of the parameters ##

plot the distribution per parameter using a tidybayes halfeye plot

```{r}
p2.1 <- ggplot(theta_group_1, aes(x = value, y = fct_rev(param))) +  
  stat_halfeye(alpha = 0.9) +
  labs(x = "parameter estimate", y = "") 
p2.1

ggsave("figures/all_params_1.jpeg",
       width = 8, height = 6)
```

plot avg and individual pid parameters

```{r}
p2.2 <- ggplot(theta_group_1, aes(x = value, y = fct_rev(param))) +  
  stat_halfeye(alpha = 0.9, position = position_nudge(y=-0.5)) +
  geom_pointinterval(data = theta_pid_1_q, aes(xmin=.lower, xmax=.upper, 
                                           colour = pid, fill = pid),
                     position = position_dodge(width = 0.7)) +
  labs(x = "parameter estimate", y = "") +
  facet_wrap(~param, scales = "free_y") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
p2.2

ggsave("figures/all_params_pid_1.jpeg",
       width = 8, height = 6)
```

# section 3 #

## plot thresholds separately ##

first, wrangle to make a separate dfs

thresholds df per pid

```{r}
thresh_pid_1 <- array_theta_pid_1 %>% 
  filter(str_detect(param, "B_")) %>% 
  mutate(response = if_else(str_detect(param, "lRLearned"), "Learned", "Novel"),
         response = factor(response,
                            levels = c("Novel", "Learned")))
head(thresh_pid_1)
str(thresh_pid_1)
summary(thresh_pid_1)

## data check
thresh_pid_1 %>% 
  distinct(param, response)
```

thresholds at the group level

```{r}
thresh_group_1 <- theta_group_1 %>% 
  filter(str_detect(param, "B_")) %>% 
  mutate(response = if_else(str_detect(param, "lRLearned"), "Learned", "Novel"),
         response = factor(response,
                            levels = c("Novel", "Learned")))

head(thresh_group_1)
str(thresh_group_1)
summary(thresh_group_1)
```

## plot ##


at the group level

```{r}
p3.1 <- ggplot(thresh_group_1, aes(x = value, y = fct_rev(response))) +
  stat_halfeye(alpha = 0.7, aes(fill = response)) +
  scale_fill_manual(values = myColors) +
  labs(y="Response", x="Threshold", title=bquote("Group level: 
Threshold by response")) +
   theme(legend.position = "none")
  
p3.1

## save a figure
 ggsave("figures/thresh.jpeg",width = 10, height = 6)
```

## caluclate a threshold response key difference ##

per pid

```{r}
thresh_diff_pid_1 <- thresh_pid_1 %>%
  pivot_wider(id_cols = -param,
              names_from = "response",
              values_from = "value") %>% 
  mutate(diff = Novel - Learned)
head(thresh_diff_pid_1)
```

at the group level

```{r}
thresh_diff_1 <- thresh_group_1 %>%
  pivot_wider(id_cols = -param,
              names_from = "response",
              values_from = "value") %>% 
  mutate(diff = Novel - Learned)
head(thresh_diff_1)
```

## quantiles on the difference score ##

per pid

```{r}
thresh_diff_pid_1_q <- thresh_diff_pid_1 %>%
  group_by(pid) %>% 
  median_qi(diff)
thresh_diff_pid_1_q
```

at the group level

```{r}
thresh_diff_1_q <- thresh_diff_1 %>%
  group_by(pid) %>%
  median_qi(diff)
thresh_diff_1_q
```

## plot the threshold difference score ##

plot the average difference

```{r}
p3.2 <- ggplot(thresh_diff_1, aes(x = diff, y = "")) +  
  stat_halfeye(alpha = 0.9, fill="firebrick") +
  geom_vline(xintercept = 0, color = "darkgrey") +
  labs(title = "Group level: 
Threshold (Novel > Learned)", 
       x = "Threshold", y="") 
p3.2
```

plot individual pids

```{r}
thresh_diff_pid_1$pid = with(thresh_diff_pid_1, reorder(pid, diff, mean))
p3.3 <- ggplot(thresh_diff_pid_1, aes(x = diff, y = pid,
                                    fill=pid)) +  
  stat_halfeye(alpha = 0.7) +
  geom_vline(xintercept = 0, color = "darkgrey") +
  labs(title = "Participant level: Threshold (Novel > Learned)", 
       x = "Threshold", y="participant") 
p3.3
```

plot the average difference and individual pids

```{r}
p3.4 <- ggplot(thresh_diff_1, aes(x = diff, y= reorder(pid, diff, mean))) +  
  stat_halfeye(data = thresh_diff_pid_1,
               alpha = 0.7) +
  geom_vline(xintercept = 0, color = "darkgrey") +
  labs(title = "Individual and average
threshold differences", 
       x = "Threshold", y="participant ID") +
  stat_halfeye(alpha = 0.9, fill="firebrick",
               position = position_nudge(y=0)) 
p3.4
```

## plot together ##

```{r}
p3.5 <- p3.1 / p3.2 | p3.3 +
  labs(caption = "80% and 95% quantile intervals shown in black")
p3.5

# save a figure
 ggsave("figures/thresh_1.jpeg",
        width = 21, height = 10)
```

and now combine avg and pid difference plots

```{r}
p3.6 <- p3.1 | p3.2 +
  labs(caption = "80% and 95% credible intervals shown in black")
p3.6

# save a figure
ggsave("figures/thresh_2.jpeg",
       width = 15, height = 10)
```

```{r}
p3.6 <- p3.1 | p3.4 +
  labs(caption = "80% and 95% quantile intervals shown in black")
p3.6

# save a figure
ggsave("figures/thresh_3.jpeg",
       width = 21, height = 10)
```


the exact layout, labels of factors and arrangement may need polishing and re-arranging.

But for now, the data seem to be what we want.

# section 4 #

## plot drift rate separately ##

## wrangle ##

create a drift-rate df

```{r}
drift_rate_pid_1 <- array_theta_pid_1 %>% 
  filter(str_detect(param, "v_S")) %>% 
  mutate(condition = if_else(str_detect(param, "Novel"), "Novel", "Learned"),
         accuracy = if_else(str_detect(param, "TRUE"), "True", "False"),
         condition = factor(condition,
                            levels = c("Novel", "Learned")),
         accuracy = factor(accuracy,
                            levels = c("False", "True")))
head(drift_rate_pid_1)
str(drift_rate_pid_1)
summary(drift_rate_pid_1)

## data check
drift_rate_pid_1 %>% 
  distinct(param, condition, accuracy)
```


drift rate at the group level

```{r}
drift_rate_group_1 <- theta_group_1 %>% 
  filter(str_detect(param, "v_S")) %>% 
  mutate(condition = if_else(str_detect(param, "Novel"), "Novel", "Learned"),
         accuracy = if_else(str_detect(param, "TRUE"), "True", "False"),
         condition = factor(condition,
                            levels = c("Novel", "Learned")),
         accuracy = factor(accuracy,
                            levels = c("False", "True")))
head(drift_rate_group_1)
str(drift_rate_group_1)
summary(drift_rate_group_1)
```

## plot ##

#Create a custom manual colour scale: drift_rate group

```{r}
library(RColorBrewer)
myColors_drift <- RColorBrewer::brewer.pal(12, "Dark2")[4:5]
names(myColors_drift) <- levels(drift_rate_group_1$accuracy)
colScale_drift <- scale_fill_manual(name = "Accuracy", values = myColors_drift)
```

the average

```{r}
p4.1 <- ggplot(drift_rate_group_1, aes(x = value, y = fct_rev(condition), 
                               fill = accuracy)) +
  stat_halfeye(alpha = 0.7) +
  colScale_drift +
  labs(title = "Group level: True and False 
drift rate by familiarity condition", y="Familiarity", x="Drift rate") 
p4.1

ggsave("figures/drift_rate_1.jpeg",
        width = 10, height = 6)
```



and now facet by pid

```{r}
p4.2 <- ggplot(drift_rate_pid_1, aes(x = value, y = fct_rev(condition), 
                               fill = accuracy)) +
  stat_halfeye(alpha = 0.7) +
  colScale_drift +
  labs(title = "Participant level: Drift-rate by condition",
       y="Familiarity", x="Drift rate") +
  facet_wrap(~pid)
p4.2
```

take a look together

```{r}
p4.1 / p4.2
ggsave("figures/drift_rate_true_false_1.jpeg", 
       width = 16, height = 13)
```

## calculate drift rate differences between conditions and accuracy ##

per pid

```{r}
## just true > false differences (accuracy)
drift_acc_diff_pid_1 <- drift_rate_pid_1 %>%
  pivot_wider(id_cols = -param,
               names_from = accuracy,
               values_from = value) %>%
  mutate(diff = True - False)
drift_acc_diff_pid_1

## unf > fam differences
drift_acc_cond_diff_pid_1 <- drift_acc_diff_pid_1 %>%
  pivot_wider(id_cols = -c("True", "False"), 
              names_from = condition,
              values_from = diff) %>%
  mutate(diff = Novel - Learned)
drift_acc_cond_diff_pid_1
```

at the group level

```{r}
## just true > false differences (accuracy)
drift_acc_diff_1 <- drift_rate_group_1 %>%
  pivot_wider(id_cols = -param,
               names_from = accuracy,
               values_from = value) %>%
  mutate(diff = True - False)
drift_acc_diff_1

## unf > fam differences
drift_acc_cond_diff_1 <- drift_acc_diff_1 %>%
  pivot_wider(id_cols = -c("True", "False"), 
              names_from = condition,
              values_from = diff) %>%
  mutate(diff = Novel - Learned)
drift_acc_cond_diff_1
```

## plot the differences ##

differences between true and false, separately for unf and fam

at the group average level


plot

```{r}
p4.3 <- ggplot(drift_acc_diff_1, aes(x = diff, y = fct_rev(condition))) +
  stat_halfeye(aes(fill = condition, alpha = 0.7)) +
  scale_fill_manual(values = myColors) +
  geom_vline(xintercept = 0, color = "darkgrey") +
  labs(title = "Group level: Drift rate (True > false)
by familiarity condition",
       x="Drift rate", y="Familiarity") +
  xlim(0,2.5) +
   theme(legend.position = "none")
  
p4.3
```

quantiles

```{r}
drift_acc_diff_1_q <- drift_acc_diff_1 %>%
  group_by(condition) %>% 
  median_qi(diff)
drift_acc_diff_1_q
```

difference between unf (true and false) > fam (true > false)

plot

```{r}
p4.4 <- ggplot(drift_acc_cond_diff_1, aes(x = diff, y = "")) +
  stat_halfeye(fill = "firebrick", alpha = 0.9) +
  geom_vline(xintercept = 0, color = "darkgrey") +
  labs(title = "Group level: Drift rate (Novel > Learned)",
       x="Drift rate", y="") +
  xlim(0,2.5)
p4.4
```

quantiles

```{r}
drift_acc_cond_diff_1_q <- drift_acc_cond_diff_1 %>%
  median_qi(diff)
drift_acc_cond_diff_1_q
```

## plot together ##

```{r}
p4.5 <- p4.3 | p4.4 +
  labs(caption = "80% and 95% credible intervals shown in black")
p4.5

# save a figure
ggsave("figures/drift_rat_mini_1.jpeg",
       width = 15, height = 10)
```

## plot the average difference and pid differences ##

```{r}
drift_acc_cond_diff_pid_1$pid = with(drift_acc_cond_diff_pid_1, reorder(pid, diff, mean))
p4.6 <- ggplot(drift_acc_cond_diff_1, aes(x = diff, y=pid)) +  
  stat_halfeye(alpha = 0.9, fill="firebrick") +
  stat_halfeye(data = drift_acc_cond_diff_pid_1,
               alpha = 0.7) +
  geom_vline(xintercept = 0, color = "darkgrey") +
  labs(title = "Individual and average 
drift rate differences", 
       x = "Drift rate", y="participant ID") 
p4.6
```

## combine drift rate plots ##

```{r}
p4.7 <- p4.3 / p4.4 | p4.6 +
  labs(caption = "80% and 95% quantile intervals shown in black")
p4.7 

# save a figure
ggsave("figures/drift_rate_pid_1.jpeg",
       width = 15, height = 10)
```

##Correlations
#load data
```{r}
data_task <- read_csv("data/data_WP3_model.csv")
data_pupil_mean <- read_csv("data/pupil_metrics/mean_pid.csv")
data_pupil_max <- read_csv("data/pupil_metrics/max_pid.csv")
data_pupil_rate <- read_csv("data/pupil_metrics/rate_pid.csv")

#Wrangle pupil data
data_pupil <- data_pupil_mean %>%
  mutate(pupil_mean = mean,
         pupil_max = data_pupil_max$mean,
         pupil_rate = data_pupil_rate$mean_pupil_change) %>%
  select(pid, condition, accuracy, pupil_mean, pupil_max, pupil_rate) %>%
  rename(cond = condition, acc = accuracy)
```

#wrangle
```{r}
data_task_sum <- data_task %>%
  select(pid,cond,acc,rt) %>%
  mutate(pid = as.factor(pid)) %>%
  group_by(pid, cond) %>%
  summarise(sum = sum(acc),
            n = n(), # n here is the total trials per grouping
            prop_correct = sum/n, 
            rt = mean(rt))
data_pupil_sum <- data_pupil %>%
  group_by(pid, cond) %>%
  summarise(pupil_mean = mean(pupil_mean),
            pupil_max = mean(pupil_max),
            pupil_rate = mean(pupil_rate))

data_pupil_corr <- data_task_sum %>%
  add_column(pupil_mean = data_pupil_sum$pupil_mean) %>%
  add_column(pupil_max = data_pupil_sum$pupil_max) %>%
  add_column(pupil_rate = data_pupil_sum$pupil_rate)
```

#Adding parameters data
```{r}
thresh_pid_1_av <- thresh_pid_1 %>%
  group_by(pid,param) %>%
  summarise(param_median = median(value)) %>%
  pivot_wider(names_from = param, values_from = param_mean)


drift_rate_pid_1_av <- drift_rate_pid_1 %>%
  group_by(pid, param) %>%
  summarise(param_median = mean(value)) %>%
  pivot_wider(names_from = param, values_from = param_mean)

subject_param <- drift_rate_pid_1_av %>%
  left_join(
    thresh_pid_1_av %>%
      select(pid, B_lRLearned, B_lRNovel), 
    by = "pid"
  )
```



## Correlations


#Correlation between pupil mean and overall accuracy during task
```{r}
overall_corr <- data_pupil_corr %>%
  group_by(pid, pupil_mean, pupil_max, pupil_rate) %>%
  summarise(accuracy = mean(prop_correct), .groups = "drop")

correlation_overall_mean <- cor(overall_corr$pupil_mean, overall_corr$accuracy)
cat("Correlation between pupil mean and overall accuracy:", round(correlation_overall_mean, 3), "\n")
```


#Correlation between pupil max and overall accuracy during task
```{r}
correlation_overall_max <- cor(overall_corr$pupil_max, overall_corr$accuracy)
cat("Correlation between maximum pupil dilation and overall accuracy:", round(correlation_overall_max, 3), "\n")
```

#Correlation between pupil rate and overall accuracy during task
```{r}
correlation_overall_rate <- cor(overall_corr$pupil_rate, overall_corr$accuracy)
cat("Correlation between rate of pupil change and overall accuracy:", round(correlation_overall_rate, 3), "\n")
```

#Plotting individual differences
```{r}
#| label: plot-individual-differences-face

p2.1 <- data_pupil_corr %>%
  ggplot(aes(x = pupil_mean, y = prop_correct, color = cond)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = TRUE) +
  facet_wrap(~cond) +
  stat_cor(method = "pearson") +
  labs(
    title = "Correlation of mean pupil size and overall accuracy",
    x = "Mean pupil size (mm)",
    y = "Proportion Correct",
    color = "Condition"
  ) + scale_color_manual(values = c("Novel" = paletteer_d("rcartocolor::Safe",4)[1],"Learned" =  paletteer_d("rcartocolor::Safe",4)[2])) 
print(p2.1)

ggsave("figures/cor/correlation_mean_condition.jpeg",
       width = 12, height =8)
```

Overall
```{r}
overall_data_pupil_corr <- data_pupil_corr %>%
  group_by(pid) %>%
  summarise(pupil_mean = mean(pupil_mean),
            pupil_max = mean(pupil_max),
            pupil_rate = mean(pupil_rate),
            prop_correct = mean(prop_correct))

p2.1.1 <- overall_data_pupil_corr %>%
  ggplot(aes(x = pupil_mean, y = prop_correct)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = TRUE) +
  stat_cor(method = "pearson") +
  labs(
    title = "Correlation between mean pupil size
and accuracy",
    x = "Mean pupil size (mm)",
    y = "Proportion Correct",
  ) 
  
print(p2.1.1)

ggsave(ggsave("figures/cor/correlation_mean_accuracy_overall.jpeg",
       width = 12, height =8))
```


#Max pupil size
#Plotting individual differences
```{r}
#| label: plot-individual-differences

p2.2 <- data_pupil_corr %>%
  ggplot(aes(x = pupil_max, y = prop_correct, color = cond)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = TRUE) +
  facet_wrap(~cond) +
  stat_cor(method = "pearson") +
  labs(
    title = "Correlation of maximum pupil dilation and overall accuracy",
    x = "Maximum pupil dilation (mm)",
    y = "Proportion Correct",
    color = "Condition"
  ) + scale_color_manual(values = c("Novel" = paletteer_d("rcartocolor::Safe",4)[1],"Learned" =  paletteer_d("rcartocolor::Safe",4)[2]))
  
print(p2.2)

ggsave("figures/cor/correlation_max_accuracy_condition.jpeg",
       width = 12, height =8)
```

Overall
```{r}
p2.2.1 <- overall_data_pupil_corr %>%
  ggplot(aes(x = pupil_max, y = prop_correct)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = TRUE) +
  stat_cor(method = "pearson") +
  labs(
    title = "Correlation between maximum pupil dilation
and accuracy",
    x = "Maximum pupil dilation (mm)",
    y = "Proportion Correct",
  ) 
  
print(p2.2.1)

ggsave(ggsave("figures/cor/correlation_max_accuracy_overall.jpeg",
       width = 12, height =8))
```

#Rate of pupil change
```{r}
p2.2 <- data_pupil_corr %>%
  ggplot(aes(x = pupil_rate, y = prop_correct, color = cond)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = TRUE) +
  facet_wrap(~cond) +
  stat_cor(method = "pearson") +
  labs(
    title = "Correlation of rate of pupil change and overall accuracy",
    x = "Rate of pupil change (mm/ms)",
    y = "Proportion Correct",
    color = "Condition"
  ) + scale_color_manual(values = c("Novel" = paletteer_d("rcartocolor::Safe",4)[1],"Learned" =  paletteer_d("rcartocolor::Safe",4)[2]))
  
print(p2.2)

ggsave(ggsave("figures/cor/correlation_rate_accuracy_condition.jpeg",
       width = 12, height =8))
```

```{r}
p2.2.1 <- overall_data_pupil_corr %>%
  ggplot(aes(x = pupil_rate, y = prop_correct)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = TRUE) +
  stat_cor(method = "pearson") +
  labs(
    title = "Correlation between rate of pupil change
and accuracy",
    x = "Rate of pupil change (mm/ms)",
    y = "Proportion Correct",
  ) 
  
print(p2.2.1)

ggsave("figures/cor/correlation_rate_accuracy_overall.jpeg",
       width = 12, height =8)
```

Correlations with parameters

```{r}
library(dplyr)
library(tidyr)
library(purrr)
library(stringr)

# Step 1: acc-specific columns (e.g. mean_learned_true)
acc_wide <- data_pupil %>%
  mutate(acc_label = ifelse(acc == 1, "true", "false"),
         cond = tolower(cond)) %>%
  group_by(pid, cond, acc_label) %>%
  summarise(across(c(pupil_mean, pupil_max, pupil_rate), mean), .groups = "drop") %>%
  pivot_wider(
    names_from = c(cond, acc_label),
    values_from = c(pupil_mean, pupil_max, pupil_rate),
    names_sep = "_"
  ) %>%
  rename_with(~ str_remove(.x, "pupil_"), starts_with("pupil_"))

# Step 2: condition-level columns collapsed across acc (e.g. mean_learned)
cond_wide <- data_pupil %>%
  mutate(cond = tolower(cond)) %>%
  group_by(pid, cond) %>%
  summarise(across(c(pupil_mean, pupil_max, pupil_rate), mean), .groups = "drop") %>%
  pivot_wider(
    names_from = cond,
    values_from = c(pupil_mean, pupil_max, pupil_rate),
    names_sep = "_"
  ) %>%
  rename_with(~ str_remove(.x, "pupil_"), starts_with("pupil_"))

# Step 3: join both
pupil_wide <- left_join(cond_wide, acc_wide, by = "pid")

# Join parameter together
thresh_wide <- thresh_pid_1 |>
  select(pid, param, value, datapoint) |>
  pivot_wider(names_from = param, values_from = value)

drift_wide <- drift_rate_pid_1 |>
  select(pid, param, value, datapoint) |>
  pivot_wider(names_from = param, values_from = value)

derived_samples_wide <- thresh_wide |>
  left_join(drift_wide, by = c("pid", "datapoint"))


# ---- 1. Prepare parameter dataframe ----
params <- param_1_pid %>%
  mutate(
    threshold_diff = B_lRNovel - B_lRLearned,
    drift_rate_learned = `v_SLearned:lMTRUE` - `v_SLearned:lMFALSE`,
    drift_rate_novel   = `v_SNovel:lMTRUE` - `v_SNovel:lMFALSE`,
    drift_diff = drift_rate_novel - drift_rate_learned
  ) %>%
  rename(
    threshold_learned = B_lRLearned,
    threshold_novel    = B_lRNovel,
    drift_learned_false = `v_SLearned:lMFALSE`,
    drift_learned_true  = `v_SLearned:lMTRUE`,
    drift_novel_false   = `v_SNovel:lMFALSE`,
    drift_novel_true    = `v_SNovel:lMTRUE`,
    pid = "subjects"
  )

# ---- 2. Add matching pupil difference scores to pupil_wide ----
pupil_wide <- pupil_wide %>%
  mutate(
    mean_diff = mean_novel - mean_learned,
    max_diff  = max_novel  - max_learned,
    rate_diff = rate_novel - rate_learned,
    mean_diff_learned = mean_learned_true - mean_learned_false,
    mean_diff_novel   = mean_novel_true   - mean_novel_false,
    mean_diff_diff    = mean_diff_novel - mean_diff_learned,
    max_diff_learned = max_learned_true - max_learned_false,
    max_diff_novel   = max_novel_true   - max_novel_false,
    max_diff_diff    = max_diff_novel - max_diff_learned,
    rate_diff_learned = rate_learned_true - rate_learned_false,
    rate_diff_novel   = rate_novel_true   - rate_novel_false,
    rate_diff_diff    = rate_diff_novel - rate_diff_learned,
  ) 


params_median <- params %>%
  group_by(pid) %>%
  summarise(across(
    c(threshold_learned, threshold_novel,
      drift_learned_true, drift_learned_false,
      drift_novel_true, drift_novel_false,
      drift_rate_learned, drift_rate_novel),
    median, .names = "{.col}"
  ), .groups = "drop")

# ---- 3. Merge ----
full_df <- left_join(params_median, pupil_wide, by = "pid")

# ---- 4. Define meaningful parameter <-> pupil metric pairings ----
# Each entry: parameter column name -> matching pupil column suffix (per metric)
pairings <- tribble(
  ~param_col,            ~pupil_suffix,
  "threshold_learned",   "learned",
  "threshold_novel",     "novel",
  "threshold_diff",      "diff",
  "drift_learned_true",  "learned_true",
  "drift_learned_false", "learned_false",
  "drift_novel_true",    "novel_true",
  "drift_novel_false",   "novel_false",
  "drift_rate_learned",  "diff_learned",
  "drift_rate_novel",    "diff_novel",
  "drift_diff",          "diff_diff"
)

metrics <- c("mean", "max", "rate")

pairings_full <- pairings %>%
  crossing(metric = metrics) %>%
  mutate(pupil_col = paste0(metric, "_", pupil_suffix)) %>%
  filter(pupil_col %in% names(pupil_wide))

# ---- 5. Compute correlations ----
cor_results <- pairings_full %>%
  mutate(
    test = map2(param_col, pupil_col, function(p, q) {
      x <- full_df[[p]]
      y <- full_df[[q]]
      ct <- cor.test(x, y, use = "pairwise.complete.obs")
      tibble(r = unname(ct$estimate), p_value = ct$p.value, n = sum(complete.cases(x, y)))
    })
  ) %>%
  unnest(test) %>%
  select(parameter = param_col, metric, pupil_column = pupil_col, r, p_value, n)

print(cor_results)




params_samples <- param_1_pid %>%
  mutate(
    threshold_diff      = B_lRNovel - B_lRLearned,
    drift_rate_learned  = `v_SLearned:lMTRUE` - `v_SLearned:lMFALSE`,
    drift_rate_novel    = `v_SNovel:lMTRUE` - `v_SNovel:lMFALSE`,
    drift_diff          = drift_rate_novel - drift_rate_learned
  ) %>%
  rename(
    threshold_learned   = B_lRLearned,
    threshold_novel     = B_lRNovel,
    drift_learned_false = `v_SLearned:lMFALSE`,
    drift_learned_true  = `v_SLearned:lMTRUE`,
    drift_novel_false   = `v_SNovel:lMFALSE`,
    drift_novel_true    = `v_SNovel:lMTRUE`,
    pid = "subjects"
  )



# Join with pupil_wide — pupil values repeat across all datapoints per pid
params_merged <- params_samples %>%
  left_join(pupil_wide, by = "pid")

params_merged <- params_samples %>%
  group_by(pid) %>%
  mutate(datapoint = row_number()) %>%
  ungroup() %>%
  left_join(pupil_wide, by = "pid")

# ---- 7. Compute r for every sample, for every pairing ----
compute_sample_cors <- function(param_col, pupil_col, data) {
  data_split <- split(data, data$datapoint)  # use datapoint, not sample
  vapply(data_split, function(d) {
    cor(d[[param_col]], d[[pupil_col]], use = "pairwise.complete.obs")
  }, numeric(1))
}

sample_cors <- pairings_full %>%
  mutate(
    r_values = map2(param_col, pupil_col, ~ compute_sample_cors(.x, .y, params_merged))
  )

# ---- 8. Unnest into long dataframe ----
cor_results_samples <- sample_cors %>%
  select(param_col, metric, pupil_col, r_values) %>%
  mutate(sample = map(r_values, ~ as.numeric(names(.x)))) %>%
  unnest(c(r_values, sample)) %>%
  rename(parameter = param_col, pupil_metric = metric, pupil_column = pupil_col, r = r_values)

print(cor_results_samples)
```


## Plotting the correlations
```{r}
library(ggplot2)

plot_data_list <- pairings_full %>%
  mutate(
    plot_df = map2(param_col, pupil_col, function(p, q) {
      full_df %>%
        select(pid, x = all_of(p), y = all_of(q)) %>%
        mutate(parameter = p, pupil_col = q)
    })
  ) %>%
  pull(plot_df) %>%
  bind_rows() %>%
  left_join(cor_results, by = c("parameter" = "parameter", "pupil_col" = "pupil_column"))

make_scatter_plot <- function(metric_name) {
  df_sub <- plot_data_list %>%
    semi_join(pairings_full %>% filter(metric == metric_name),
              by = c("parameter" = "param_col", "pupil_col" = "pupil_col"))

  ggplot(df_sub, aes(x = x, y = y)) +
    geom_point(size = 2, alpha = 0.7) +
    geom_smooth(method = "lm", se = TRUE, color = "steelblue") +
    facet_wrap(~ parameter, scales = "free") +
    geom_text(
      data = df_sub %>% distinct(parameter, r, p_value),
      aes(x = -Inf, y = Inf,
          label = sprintf("r = %.2f, p = %.3f", r, p_value)),
      hjust = -0.1, vjust = 1.5, inherit.aes = FALSE, size = 3.5
    ) +
    labs(title = paste("Correlations with pupil", metric_name, "(median parameter values)"),
         x = "LBA parameter (median posterior value)", y = paste("Pupil", metric_name)) +
    theme_minimal()
}

plot_mean <- make_scatter_plot("mean")
plot_max  <- make_scatter_plot("max")
plot_rate <- make_scatter_plot("rate")

plot_mean
ggsave("figures/cor/point_estimate_cor_mean.jpeg", height = 10, width = 15)
plot_max
ggsave("figures/cor/point_estimate_cor_max.jpeg", height = 10, width = 15)
plot_rate
ggsave("figures/cor/point_estimate_cor_rate.jpeg", height = 10, width = 15)
```

```{r}
library(ggdist)

make_halfeye_plot <- function(metric_name) {
  df_sub <- cor_results_samples %>%
    filter(pupil_metric == metric_name)
  
  ggplot(df_sub, aes(x = r, y = parameter)) +
    stat_halfeye(fill = "steelblue", alpha = 0.7) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
    facet_wrap(~ parameter, scales = "free_y", ncol = 1) +
    labs(title = paste("Posterior correlations with pupil", metric_name),
         x = "Correlation (r)", y = NULL) +
    theme_minimal() +
    theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())
}

halfeye_mean <- make_halfeye_plot("mean")
halfeye_max  <- make_halfeye_plot("max")
halfeye_rate <- make_halfeye_plot("rate")

halfeye_mean
ggsave("figures/cor/post_cor_mean.jpeg", height = 12, width = 10)
halfeye_max
ggsave("figures/cor/post_cor_max.jpeg", height = 12, width = 10)
halfeye_rate
ggsave("figures/cor/post_cor_rate.jpeg", height = 12, width = 10)
```


Plotting contrasts:
```{r}
library(paletteer)
library(patchwork)

myColors <- c("Novel" = paletteer_d("rcartocolor::Safe", 4)[1],
              "Learned" = paletteer_d("rcartocolor::Safe", 4)[2])

make_halfeye_contrast_plot <- function(metric_name) {
  
  # Drift rate contrast: learned vs novel
  drift_contrast <- cor_results_samples %>%
    filter(pupil_metric == metric_name,
           parameter %in% c("drift_rate_learned", "drift_rate_novel")) %>%
    mutate(
      condition = ifelse(parameter == "drift_rate_novel", "Novel", "Learned"),
      contrast  = "Drift rate: Learned vs Novel"
    )
  
  # Threshold contrast: learned vs novel
  thresh_contrast <- cor_results_samples %>%
    filter(pupil_metric == metric_name,
           parameter %in% c("threshold_learned", "threshold_novel")) %>%
    mutate(
      condition = ifelse(parameter == "threshold_novel", "Novel", "Learned"),
      contrast  = "Threshold: Learned vs Novel"
    )
  
  df_sub <- bind_rows(drift_contrast, thresh_contrast)
  
  ggplot(df_sub, aes(x = r, y = condition, fill = condition)) +
    stat_halfeye(alpha = 0.7) +
    scale_fill_manual(values = myColors) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
    facet_wrap(~ contrast, ncol = 1, scales = "free_y") +
    labs(
      title = paste("Posterior correlation contrasts — pupil", metric_name),
      x = "Correlation (r)", y = NULL, fill = "Condition"
    ) +
    theme_minimal() +
    theme(
      text = element_text(face = "bold"),          
      axis.text.y = element_text(face = "bold"),   
      axis.ticks.y = element_blank(),
      legend.position = "none"
)
}

contrast_mean <- make_halfeye_contrast_plot("mean")
contrast_max  <- make_halfeye_contrast_plot("max")
contrast_rate <- make_halfeye_contrast_plot("rate")

contrast_mean
ggsave("figures/cor/post_cor_mean_contrasts.jpeg", height = 12, width = 10)
contrast_max
ggsave("figures/cor/post_cor_max_contrasts.jpeg", height = 12, width = 10)
contrast_rate
ggsave("figures/cor/post_cor_rate_contrasts.jpeg", height = 12, width = 10)



```






PLot posterior correlation distributions
Mean Pupil

Disclaimer: here only the correlations that make sense (Learned_Parameter with Learned_pupil metrics)
```{r}
# Plot face
p10.1 <- ggplot(correlation_draws_rate, aes(x = correlation_rate_learned_true)) +  
  stat_halfeye(alpha = 0.9, .width = c(0.5, 0.95)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Posterior Correlation Distributions",
    subtitle = "Correlations between Mean pupil size and LBA parameters across draws",
    x = "Correlation with Face Recognition", 
    y = "Density"
  ) +
  facet_wrap(~parameter, scale = "free_x", nrow = 2) 

print(p10.1)

ggsave("figures/posterior_correlation_distribution_mean_learned.jpeg",
       width = 16, height = 12)

# Plot car
p10.2 <- ggplot(correlation_draws, aes(x = correlation_car)) +  
  stat_halfeye(alpha = 0.9, .width = c(0.5, 0.95)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Posterior Correlation Distributions",
    subtitle = "Correlations between car recognition and LBA parameters across draws",
    x = "Correlation with Car Recognition", 
    y = "Density"
  ) +
  facet_wrap(~parameter, scale = "free_x", nrow = 1) 

print(p10.2)

ggsave("WP2_1/figures/posterior_correlation_distribution_car.jpeg",
       width = 16, height = 12)


```



Plot threshold and drift rate separately
# Mean pupil size
```{r}
#Plot mean / threshold
facets <- c("B_lRNovel","B_lRLearned", "threshold_diff")
p10.3 <- ggplot(correlation_draws[correlation_draws$parameter %in% facets,], aes(x = correlation_face)) +  
  stat_halfeye(alpha = 0.9, .width = c(0.5, 0.95)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Posterior Correlation Distributions",
    subtitle = "Correlations between face recognition and threshold parameters across draws",
    x = "Correlation with Face Recognition", 
    y = "Density"
  ) +
  facet_wrap(~factor(parameter, facets), scale = "free_x", nrow = 1) 

print(p10.3)

ggsave("WP2_1/figures/posterior_correlation_distribution_face_threshold.jpeg",
       width = 16, height = 12)

#Plot face drift rate
facets <- c("v_SLearned:lMFALSE","v_SLearned:lMTRUE", "drift_diff_learned","v_SNovel:lMFALSE", "v_SNovel:lMTRUE","drift_diff_novel")
p10.4 <- ggplot(correlation_draws[correlation_draws$parameter %in% facets,], aes(x = correlation_face)) +  
  stat_halfeye(alpha = 0.9, .width = c(0.5, 0.95)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Posterior Correlation Distributions",
    subtitle = "Correlations between face recognition and drift rate parameters across draws",
    x = "Correlation with Face Recognition", 
    y = "Density"
  ) +
  facet_wrap(~factor(parameter, facets), scale = "free_x", nrow = 2) 

print(p10.4)

ggsave("WP2_1/figures/posterior_correlation_distribution_face_drift.jpeg",
       width = 16, height = 12)

```
#Car
```{r}
#Plot car threshold
facets <- c("B_lRNovel","B_lRLearned", "threshold_diff")
p10.3 <- ggplot(correlation_draws[correlation_draws$parameter %in% facets,], aes(x = correlation_car)) +  
  stat_halfeye(alpha = 0.9, .width = c(0.5, 0.95)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Posterior Correlation Distributions",
    subtitle = "Correlations between face recognition and threshold parameters across draws",
    x = "Correlation with Face Recognition", 
    y = "Density"
  ) +
  facet_wrap(~factor(parameter, facets), scale = "free_x", nrow = 1) 

print(p10.3)

ggsave("WP2_1/figures/posterior_correlation_distribution_face_threshold.jpeg",
       width = 16, height = 12)

#Plot car drift rate
facets <- c("v_SLearned:lMFALSE","v_SLearned:lMTRUE", "drift_diff_learned","v_SNovel:lMFALSE", "v_SNovel:lMTRUE","drift_diff_novel")
p10.4 <- ggplot(correlation_draws[correlation_draws$parameter %in% facets,], aes(x = correlation_car)) +  
  stat_halfeye(alpha = 0.9, .width = c(0.5, 0.95)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Posterior Correlation Distributions",
    subtitle = "Correlations between face recognition and drift rate parameters across draws",
    x = "Correlation with Face Recognition", 
    y = "Density"
  ) +
  facet_wrap(~factor(parameter, facets), scale = "free_x", nrow = 2) 

print(p10.4)

ggsave("WP2_1/figures/posterior_correlation_distribution_face_drift.jpeg",
       width = 16, height = 12)

```
Plot together correlations for face and car
```{r}
correlation_draws_plot <- correlation_draws %>%
  pivot_longer(!c(sample, parameter, n_subjects), names_to = "Recognition_task", values_to = "values") %>%
  mutate(Recognition_task = as.factor(Recognition_task),
    Recognition_task = fct_recode(Recognition_task,"Car score correlation" = "correlation_car", "Face score correlation" = "correlation_face"))

#Plots for threshold
facets_thresh <- c("B_lRNovel","B_lRLearned", "threshold_diff")
p10.5 <- ggplot(correlation_draws_plot[correlation_draws_plot$parameter %in% facets_thresh,], aes(x = values)) +  
  stat_halfeye(aes(fill = Recognition_task), alpha = 0.9, .width = c(0.5, 0.95)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Posterior Correlation Distributions",
    subtitle = "Correlations between face recognition and car recognition 
with threshold parameters across draws",
    x = "Correlation", 
    y = "Density"
  ) +
  facet_wrap(~factor(parameter, facets_thresh), scale = "free_x", nrow = 1) 

print(p10.5)

ggsave("WP2_1/figures/posterior_correlation_distribution_car_face_thresh.jpeg",
       width = 16, height = 12)

#Plots for drift rate
facets_drift <- c("v_SLearned:lMFALSE","v_SLearned:lMTRUE", "drift_diff_learned","v_SNovel:lMFALSE", "v_SNovel:lMTRUE","drift_diff_novel")
p10.6 <- ggplot(correlation_draws_plot[correlation_draws_plot$parameter %in% facets_drift,], aes(x = values)) +  
  stat_halfeye(aes(fill = Recognition_task), alpha = 0.9, .width = c(0.5, 0.95)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Posterior Correlation Distributions",
    subtitle = "Correlations between face recognition and car recognition
and drift rate parameters across draws",
    x = "Correlation", 
    y = "Density"
  ) +
  facet_wrap(~factor(parameter, facets_drift), scale = "free_x", nrow = 2) 

print(p10.6)

ggsave("WP2_1/figures/posterior_correlation_distribution_car_face_drift.jpeg",
       width = 16, height = 12)
```
Difference between car and face posterior correlations
```{r}
#Plots for threshold
correlations_diff_thresh <- correlation_draws_plot %>%
  filter(parameter == "B_lRLearned" | parameter == "B_lRNovel" |parameter == "threshold_diff") %>%
  group_by(sample, parameter) %>%
  summarise(diff = values[Recognition_task == "Face score correlation"] - values[Recognition_task == "Car score correlation"])

facets_thresh <- c("B_lRNovel","B_lRLearned", "threshold_diff")
p10.7 <- ggplot(correlations_diff_thresh[correlations_diff_thresh$parameter %in% facets_thresh,], aes(x = diff)) +  
  stat_halfeye(fill = "firebrick", alpha = 0.9, .width = c(0.5, 0.95)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Posterior Correlation Distributions",
    subtitle = "Correlations difference between face recognition and car recognition 
with threshold parameters across draws",
    x = "Correlation difference (face > car)", 
    y = "Density"
  ) +
  facet_wrap(~factor(parameter, facets_thresh), scale = "free_x", nrow = 1) 

print(p10.7)

ggsave("WP2_1/figures/posterior_correlation_distribution_diff_thresh.jpeg",
       width = 16, height = 12)

#Plots for drift rate
correlations_diff_drift <- correlation_draws_plot %>%
  filter(parameter == "v_SLearned:lMFALSE" | parameter == "v_SLearned:lMTRUE" | parameter == "drift_diff_learned"| parameter == "v_SNovel:lMFALSE"
         | parameter == "v_SNovel:lMTRUE" | parameter == "drift_diff_novel") %>%
  group_by(sample, parameter) %>%
  summarise(diff = values[Recognition_task == "Face score correlation"] - values[Recognition_task == "Car score correlation"])

facets_drift <- c("v_SLearned:lMFALSE","v_SLearned:lMTRUE", "drift_diff_learned","v_SNovel:lMFALSE", "v_SNovel:lMTRUE","drift_diff_novel")
p10.8 <- ggplot(correlations_diff_drift[correlations_diff_drift$parameter %in% facets_drift,], aes(x = diff)) +  
  stat_halfeye(fill = "firebrick", alpha = 0.9, .width = c(0.5, 0.95)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Posterior Correlation Distributions",
    subtitle = "Correlations between face recognition and car recognition
and drift rate parameters across draws",
    x = "Correlation difference (face > car)", 
    y = "Density"
  ) +
  facet_wrap(~factor(parameter, facets_drift), scale = "free_x", nrow = 2) 

print(p10.8)

ggsave("WP2_1/figures/posterior_correlation_distribution_diff_drift.jpeg",
       width = 16, height = 12)
```



```{r}
# Check how many parameters have sufficient variability
#Face
sufficient_var_count_face <- sum(correlation_summary_face$sufficient_variability)
total_params_face <- nrow(correlation_summary_face)

cat("\nValidation Results:\n")
cat("Parameters with sufficient correlation variability (SD > 0.01):", sufficient_var_count_face, "out of", total_params_face, "\n")

#Car
sufficient_var_count_car <- sum(correlation_summary_car$sufficient_variability)
total_params_car <- nrow(correlation_summary_car)

cat("\nValidation Results:\n")
cat("Parameters with sufficient correlation variability (SD > 0.01):", sufficient_var_count_car, "out of", total_params_car, "\n")
```

#Save data
```{r}
#Save the samples for all parameters, for each participant with the face recognition score centered
write_csv(derived_samples_long, "WP2_1/data/derived_samples_long.csv")
write_csv(derived_samples_wide, "WP2_1/data/derived_samples_wide.csv")

#Save the correlations draws
write_csv(correlation_draws, "WP2_1/data/correlations_draws.csv")
write_csv(correlation_summary_car, "WP2_1/data/correlations_summary_car.csv")
write_csv(correlation_summary_face, "WP2_1/data/correlations_summary_face.csv")
```






