---
title: "modelling"
author: "Fournier Raphaël"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load the libraries that we will be using

## load packages

```{r load-pkg}
pkxg <- c("EMC2", "tidyverse", "RColorBrewer", "patchwork", "here", "future", "parallel")

lapply(pkxg, library, character.only = TRUE)
```

## plot settings

theme settings for ggplot

```{r, eval = F}
theme_set(
  theme_bw() +
    theme(text = element_text(size = 18, face="bold"), 
          title = element_text(size = 18, face="bold"),
          legend.position = "bottom")
)

## Set the amount of dodge in figures
pd <- position_dodge(0.7)
pd2 <- position_dodge(1)
```

# 1. read in the raw data and wrangle

## read in the wrangled data

Load data (30 pids)

```{r}
raw <- read_csv(here("data", "data_WP3_model.csv"))
head(raw)
str(raw)

options(mc.cores = parallel::detectCores(),
        future.fork.enable = TRUE,
        future.rng.onMisuse = "ignore") ## automatically set in RStudio

supportsMulticore()

detectCores()

plan(multicore)
```

#Wrangle

```{r}
data_emc <- raw %>%
  select(pid, cond, acc, rt) %>% 
  rename(subjects = pid, S = cond, R = acc) %>% 
  mutate(subjects=factor(subjects),
         S=factor(S,
                  levels = c("Novel", "Learned")),
         R=case_when(
           R == "1" & S == "Novel" ~ "Novel",
           R == "1" & S == "Learned" ~ "Learned",
           R == "0" & S == "Novel" ~ "Learned",
           R == "0" & S == "Learned" ~ "Novel"),
         R=factor(R,
                  levels = c("Novel", "Learned"))) %>% 
  as.data.frame() ## in DMC, post predict need a df and not a tibble. Same for EMC2 I think.
head(data_emc)
str(data_emc)
```

# 2. build model

# 2.1 Prergistered model

## set the match factor

```{r}
matchfun=function(d)d$S==d$lR
```

```{r}
design_model_wp3 <- design(data = data_emc,model=LBA,matchfun=matchfun,
                       formula=list(v~0+S:lM,sv~0+lM,B~0+lR,A~1,t0~1),
                       constants=c(sv_lMFALSE=log(1)))
```

```{r}
p_vector <- c("v_SNovel:lMFALSE"=4,"v_SLearned:lMFALSE"=4, "v_SNovel:lMTRUE"=4, "v_SLearned:lMTRUE"=4,
              sv_lMTRUE=log(.5), 
              B_lRNovel=log(2), B_lRLearned=log(2),
              A=log(.5),t0=log(.3))

mapped_pars(p_vector, x=design_model_wp3)
```

## set priors

try to copy dmc priors.

I made the A and t0 sd priors smaller as they are fixed values and it seemed odd to sample a wide range when they are fixed at .5 and .3, respectively.

```{r}
mu_mean=c(`v_SNovel:lMFALSE`=0,`v_SLearned:lMFALSE`=0, `v_SNovel:lMTRUE`=1, `v_SLearned:lMTRUE`=1,
          sv_lMTRUE=log(.5), 
          B_lRNovel=log(1), B_lRLearned=log(1),
          A=log(.5),t0=log(.3))

mu_sd=c(`v_SNovel:lMFALSE`=2,`v_SLearned:lMFALSE`=2, `v_SNovel:lMTRUE`=2, `v_SLearned:lMTRUE`=2,
          sv_lMTRUE=.5, 
          B_lRNovel=.5, B_lRLearned=.5,
          A=.3,t0=.3)

prior_wp3 <- prior(design_model_wp3, type ='standard',pmean=mu_mean,psd=mu_sd)

plot(x = prior_wp3,layout=c(3,3))
```

## fit the model




```{r}
LBA_WP3 <- make_emc(data_emc,design_model_wp3,type="standard",  prior=prior_wp3, compress = TRUE)
LBA_WP3 <- fit(LBA_WP3, cores_per_chain = 1, cores_for_chains = 1, fileName="models/tmp.RData",iter = 5000, verbose = TRUE, step_size = 100)
save(LBA_WP3,file="models/LBA_WP3.RData") 
load("models/LBA_WP3.RData")
```

# check the model

convergence efficiency?

```{r}
check(LBA_WP3)
get_pars(LBA_WP3, map = TRUE)
```

plot pars

```{r}
plot_pars(LBA_WP3,selection="mu",layout=c(2,5))
```

# calculate and plot posterior predictions

```{r}
pp_LBA_WP3 <- predict(LBA_WP3, n_cores = 10)
save(pp_LBA_WP3,file="models/pp_LBA_WP3.RData")
```

```{r}
load("models/pp_LBA_WP3.RData")
```


## 2.3 Check model

```{r}
#| label: check

check(LBA_WP3)
summary(LBA_WP3)

## save out as necessary
png("figures/lba/chains.png", width = 8, height = 6, units = "in", res = 150)
plot(
  LBA_WP3,
  selection = "mu",
  layout = c(3, 3)
)
dev.off()

```

```{r}
#| label: plot-pars

plot_pars(LBA_WP3, selection = "mu", layout = c(3, 3))

## save out as necessary
png("figures/lba/pars.png", width = 8, height = 6, units = "in", res = 300)
plot_pars(
  LBA_WP3,
  selection = "mu",
  layout = c(3, 3)
)
dev.off()
```


```{r}
#| label: plot-fit

png("figures/pp_density.png", width = 8, height = 6, units = "in", res = 150)
plot_density(
  data_emc, 
  pp_LBA_WP3, 
  factors = "S"
)
dev.off()

png("figures/pp_cdf.png", width = 8, height = 6, units = "in", res = 150)
plot_cdf(
  data_emc,
  pp_LBA_WP3,  
  factors = "S"
)
dev.off()
```

